Plots

### 2021-07-27 Abstract.Rmd
changes to plots

cos.sim.auc.outliers <- cos.sim.auc.data %>%
  filter(residuals >= quantile(cos.sim.auc.data$residuals, 0.95) | 
         residuals <= quantile(cos.sim.auc.data$residuals, 0.05))
```

@@

for (x in seq(10, num.mut-1, by = 10)){
  sample.ffpe <- create_signature_sample_vector(ffpe.sig, x)
  ffpe.truth <- rep("FFPE", x)
  sample.sig6 <- create_signature_sample_vector(cosmic.sig6, num.mut-x)
  sig6.truth <- rep("MMR", num.mut-x)
  test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
                                list(cosmic.sig6, ffpe.sig))
  classification <- test %>%
    mutate(trueBinary = ifelse(truth == "MMR",1,0))

@@

for (x in seq(10, num.mut-1, by = 10)){
  sample.ffpe <- create_signature_sample_vector(ffpe.sig, x)
  ffpe.truth <- rep("FFPE", x)
  sample.sig4 <- create_signature_sample_vector(cosmic.sig4, num.mut-x)
  sig4.truth <- rep("Tobacco", num.mut-x)
  test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"),
                                list(cosmic.sig4, ffpe.sig))
  classification <- test %>%
    mutate(trueBinary = ifelse(truth == "Tobacco",1,0))

@@

for (iteration in 1:50){
  sens.MMR <- performance(test.MMR, measure = "sens")

@@

for(iteration in 1:50){
  sens.tob <- performance(test.tob, measure = "sens")

@@

```{r}
sensitivity = append(sens.MMR.vec, sens.tob.vec)
signature.sens = append(rep("MMR", length(sens.MMR.vec)),
                        rep("Tobacco", length(sens.tob.vec)))
dilution = append(1:99, 1:99)
sens.df <- data.frame(dilution, sensitivity, signature.sens)

ggplot(data = sens.df, mapping = aes(x = dilution, 
                                     y = sensitivity, 
                                     color = signature.sens)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Sensitivity as a function of FFPE noise",
       x = "Percent of FFPE noise in samples",
       y = "Sensitivity value",
       color = "True signal")
```

@@

```{r}
specificity = append(spec.MMR.vec, spec.tob.vec)
signature.spec = append(rep("MMR", length(sens.MMR.vec)),
                        rep("Tobacco", length(spec.tob.vec)))
dilution = append(1:99, 1:99)
spec.df <- data.frame(dilution, specificity, signature.spec)

ggplot(data = spec.df, mapping = aes(x = dilution, 
                                     y = specificity, 
                                     color = signature.sens)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Specificity as a function of FFPE noise",
       x = "Percent of FFPE noise in samples",
       y = "Specificity value",
       color = "True signal")
```

>>>>

cos.sim.auc.outliers <- cos.sim.auc.data %>%
  filter(residuals >= quantile(cos.sim.auc.data$residuals, 0.90) | 
         residuals <= quantile(cos.sim.auc.data$residuals, 0.10) |
           signature %in% c("SBS4"))
```

@@

for (x in seq(10, num.mut-1, by = 10)){
  sample.sig6 <- create_signature_sample_vector(cosmic.sig6, x)
  sig6.truth <- rep("MMR", x)
  sample.ffpe <- create_signature_sample_vector(ffpe.sig, num.mut-x)
  ffpe.truth <- rep("FFPE", num.mut-x)
  test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
                                list(cosmic.sig6, ffpe.sig))
  classification <- test %>%
    mutate(trueBinary = ifelse(truth == "MMR",1,0))

@@

for (x in seq(10, num.mut-1, by = 10)){
  sample.sig4 <- create_signature_sample_vector(cosmic.sig4, x)
  sig4.truth <- rep("Tobacco", x)
  sample.ffpe <- create_signature_sample_vector(ffpe.sig, num.mut-x)
  ffpe.truth <- rep("FFPE", num.mut-x)
  test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"),
                                list(cosmic.sig4, ffpe.sig))
  classification <- test %>%
    mutate(trueBinary = ifelse(truth == "Tobacco",1,0))

@@

for (iteration in 1:20){
  sens.MMR <- performance(test.MMR, measure = "sens")

@@

for(iteration in 1:20){
  sens.tob <- performance(test.tob, measure = "sens")

@@

```{r}
plot(1:99, sens.MMR.vec, ylim = c(0,1), xlab = "Percent of each signature", 
     ylab = "Sensitivity value", type = "l", main = "Comparing performance characteristics")
lines(1:99, sens.tob.vec, lty = 2)
abline(h = 0.85, col = "red")
legend("bottomright", c("MMR sensitivity", "Tobacco sensitivity", 
                        "Satisfactory Sensitivity"),
       lty = c(1,2,1), col = c("black", "black", "red"))
```

@@

```{r}
plot(1:99, spec.MMR.vec, ylim = c(0,1), xlab = "Percent of each signature", 
     ylab = "Specificity value", type = "l", main = "Comparing performance characteristics")
lines(1:99, spec.tob.vec, lty = 2)
abline(h = 0.85, col = "red")
legend("bottomleft", c("MMR specificity", "Tobacco specificity", 
                       "Satisfactory Specificity"),
       lty = c(1,2,1), col = c("black", "black", "red"))
```

### 2021-07-27 plots.Rmd
changes to plots

#Scatterplot of Average AUC vs. Cosine Similarity to FFPE
```{r}
#outer for loop: loop through each COSMIC v3 signature
average.auc.values <- c()
for (x in 1:60){
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df.1)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df.1 %>%
      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
    predictions <- classification.df.2[3]
    labels <- classification.df.2$signature.indicator
    pred <- prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
  }
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

@@

cos.sim.auc.outliers <- cos.sim.auc.data %>%
  filter(residuals >= quantile(cos.sim.auc.data$residuals, 0.95) | 
         residuals <= quantile(cos.sim.auc.data$residuals, 0.05))
```

```{r}
ggplot(cos.sim.auc.data, aes(x = cosine_similarity, y = average_auc))+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title = "Average area under curve vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average area under curve (10 iterations)")+
  geom_point(data = cos.sim.auc.outliers, color = "red")+
  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)

@@

@@

```{r}
#outer for loop: loop through each COSMIC v3 signature
average.sensitivity.values <- c()
average.specificity.values <- c()
for (x in 1:60){
  signature.sensitivity.vec <- c()
  signature.specificity.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df)
    sig.name <- class.df.colnames[3]
    #Count number of true positives (for sensititivity)
    sensitivity.df <- classification.df %>%
      filter(truth == sig.name & classify == sig.name)
    #Calculate sensitivity
    sensitivity.val = nrow(sensitivity.df)/100
    #Add to the vector of 10 sensitivity values
    signature.sensitivity.vec[y] = sensitivity.val
    #Count number of true negatives (for specificity)
    specificity.df <- classification.df %>%
      filter(truth == "FFPE" & classify == "FFPE")
    #Calculate specificity
    specificity.val = nrow(specificity.df)/100
    #Add to the vector of 10 specificity values
    signature.specificity.vec[y] = specificity.val
  }
  average.signature.sensitivity <- mean(signature.sensitivity.vec)
  average.sensitivity.values[x] <- average.signature.sensitivity
  average.signature.specificity <- mean(signature.specificity.vec)
  average.specificity.values[x] <- average.signature.specificity
}
```

@@

```{r}
ggplot(data = sens.spec.data.2, aes(x = cosine_similarity, y = value, color = measure))+
  geom_point()+
  geom_smooth(method = lm)+
  labs(title = "Sensitivity & Specificity vs. Cosine Similarity to FFPE Signature", x = "Cosine Similarity of COSMIC Signature to FFPE", y = "Sensitivity and Specificity")
```

>>>>

#Scatterplot of Average AUC vs. Cosine Similarity to FFPE
```{r}
#outer for loop: loop through each COSMIC v3 signature
average.auc.values <- c()
for (x in 1:60){
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],500)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 500)
    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df.1)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df.1 %>%
      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
    predictions <- classification.df.2[3]
    labels <- classification.df.2$signature.indicator
    pred <- prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
  }
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

@@

cos.sim.auc.outliers <- cos.sim.auc.data %>%
  filter(residuals >= quantile(cos.sim.auc.data$residuals, 0.90) | 
         residuals <= quantile(cos.sim.auc.data$residuals, 0.10) | 
         signature %in% c("SBS4"))
```

```{r}
auc.scatter <- ggplot(cos.sim.auc.data, aes(x = cosine_similarity, y = average_auc))+
  geom_point()+
  geom_smooth(method = "lm", color = mycolor[2], se = FALSE)+
  labs(title = "Average Area Under Curve vs. Cosine Similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average area under curve (10 iterations)")+
  geom_point(data = cos.sim.auc.outliers, color = mycolor [1])+
  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
  theme(plot.title = element_text(hjust = 0.5))

ggsave("aucscatter.png", plot = auc.scatter, width = 175, height = 108, units = "mm")

@@

```{r}
cos.sim.histogram <- ggplot(data = cos.sim.auc.data, aes(x = cosine_similarity))+
  geom_histogram(fill = mycolor[2], bins = 20)+
  labs(title = "Distribution of Cosine Simlarities Between COSMIC Signatures and FFPE", x = "Cosine Similarity", y = "Count")+
  theme(plot.title = element_text(hjust = 0.5))

ggsave("cossimhistogram.png", plot = cos.sim.histogram, width = 175, height = 108, units = "mm")
```

```{r}
auc.histogram <- ggplot(data = cos.sim.auc.data, aes(x = average_auc))+
  geom_histogram(fill = mycolor[2], bins = 20)+
  labs(title = "Distribution of Average AUC Values", x = "Average AUC", y = "Count")+
  theme(plot.title = element_text(hjust = 0.5))

ggsave("auchistogram.png", plot = auc.histogram, width = 175, height = 108, units = "mm")
```

@@

```{r}
#outer for loop: loop through each COSMIC v3 signature
average.sensitivity.values <- c()
average.specificity.values <- c()
for (x in 1:60){
  signature.sensitivity.vec <- c()
  signature.specificity.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],500)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 500)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df)
    sig.name <- class.df.colnames[3]
    #Count number of true positives (for sensititivity)
    sensitivity.df <- classification.df %>%
      filter(truth == sig.name & classify == sig.name)
    #Calculate sensitivity
    sensitivity.val = nrow(sensitivity.df)/500
    #Add to the vector of 10 sensitivity values
    signature.sensitivity.vec[y] = sensitivity.val
    #Count number of true negatives (for specificity)
    specificity.df <- classification.df %>%
      filter(truth == "FFPE" & classify == "FFPE")
    #Calculate specificity
    specificity.val = nrow(specificity.df)/500
    #Add to the vector of 10 specificity values
    signature.specificity.vec[y] = specificity.val
  }
  average.signature.sensitivity <- mean(signature.sensitivity.vec)
  average.sensitivity.values[x] <- average.signature.sensitivity
  average.signature.specificity <- mean(signature.specificity.vec)
  average.specificity.values[x] <- average.signature.specificity
}
```

@@

```{r}
ggplot(data = sens.spec.data.2, aes(x = cosine_similarity, y = value, color = measure))+
  geom_point()+
  geom_smooth(method = lm, se = FALSE)+
  labs(title = "Sensitivity & Specificity vs. Cosine Similarity to FFPE Signature", x = "Cosine Similarity of COSMIC Signature to FFPE", y = "Average Sensitivity and Specificity (10 iterations)")+
  scale_color_brewer(palette = "Set1")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
mycolor <- brewer.pal(3, "Set1")

sensitivity.histogram <- ggplot(data = sens.spec.data, aes(x = sensitivity))+
  geom_histogram(fill = mycolor[2], bins = 20)+
  labs(title = "Distribution of Average Sensitivities", x = "Sensitivity", y = "Count")+
  theme(plot.title = element_text(hjust = 0.5))

ggsave("sensitivityhistogram.png", plot = sensitivity.histogram, width = 175, height = 108, units = "mm")
```

```{r}
specificity.histogram <- ggplot(data = sens.spec.data, aes(x = specificity))+
  geom_histogram(fill = mycolor[2], bins = 20)+
  labs(title = "Distribution of Average Specificities", x = "Specificity", y = "Count")+
  theme(plot.title = element_text(hjust = 0.5))

ggsave("specificityhistogram.png", plot = specificity.histogram, width = 175, height = 108, units = "mm")
```

```{r}
sens.spec.label.points <- sens.spec.data %>%
  filter(signature %in% c("SBS38", "SBS4", "SBS89", "SBS3", "SBS33", "SBS44", "SBS31", "SBS40", "SBS42", "SBS23", "SBS19", "SBS11", "SBS30"))
```

```{r}
sensitivity.scatter <- ggplot(data = sens.spec.data, mapping = aes(x = cosine_similarity, y = sensitivity))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = mycolor[2])+
  labs(title = "Average Sensitivity vs. Cosine Similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average sensitivity (10 iterations)")+
  geom_point(data = sens.spec.label.points, color = mycolor [1])+
  geom_label_repel(data = sens.spec.label.points, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
  theme(plot.title = element_text(hjust = 0.5))

sensitivity.scatter

ggsave("sensitivityscatter.png", plot = sensitivity.scatter, width = 175, height = 108, units = "mm")
```

```{r}
specificity.scatter <- ggplot(data = sens.spec.data, mapping = aes(x = cosine_similarity, y = specificity))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = mycolor[2])+
  labs(title = "Average Specificity vs. Cosine Similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average specificity (10 iterations)")+
  geom_point(data = sens.spec.label.points, color = mycolor [1])+
  geom_label_repel(data = sens.spec.label.points, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
  theme(plot.title = element_text(hjust = 0.5))

specificity.scatter

ggsave("specificityscatter.png", plot = specificity.scatter, width = 175, height = 108, units = "mm")
```
